---
title: "Logistic regression and classifications"
editor_options: 
  chunk_output_type: console
---

```{r include=FALSE}
knitr::opts_chunk$set(message=FALSE)
```

```{r, echo=FALSE}
htmltools::includeHTML("questions/openai-api.html")

```

# Logistic regression and classifications

Please read the lecture to gain the understanding of logistic regression.

# Application 1: Mortgage loan
The Great Recession has forced financial institutions to be extra stringent in granting mortgage loans. Thirty recent mortgage applications are obtained to analyze the mortgage approval rate.

The response variable $y$ equals 1 if the mortgage loan is approved, 0 otherwise.
It is believed that approval depends on the percentage of the down payment ($x_1$) and the percentage of the income-to-loan ratio ($x_2$).

## Download and import the [`Morgage.csv`](examples/Morgage.csv) dataset

```{r, message=FALSE}
library(readr)
Morgage <- read_csv("examples/Morgage.csv")
```

```{r, echo=FALSE}
library(kableExtra)
Morgage %>%
  kbl() %>%
  kable_paper(full_width = F, position = "center") %>%
  scroll_box(width = "500px", height = "200px")
```

## Fit the linear probability model

Try to ask EducateUsGPT assistant to figure out the codes that can help you to develop the linear probability model for the mortgage dataset. 

```{r, echo=FALSE, results='asis'}
# Me: How to fit a linear probability model in R for the dataset that has y as response, x1 and x2 as predictors?
htmltools::includeHTML("questions/Q-321-LPM.html")
```

```{r, echo=FALSE, results='hide'}
# Assuming your dataset is named 'data'
LPM_mod <- lm(y ~ x1 + x2, data = Morgage)

# Print the summary of the model
summary(LPM_mod)
```

## Fit the logistic regression model

Try to ask EducateUsGPT assistant to figure out the codes that can help you to develop the logistic regression model for the mortgage dataset. 

```{r, echo=FALSE, results='asis'}
# Me: What question you should ask?
htmltools::includeHTML("questions/Q-322-Logit.html")
```

```{r, echo=FALSE, results='hide'}
# Assuming your dataset is named 'Morgage'
Logit_mod <- glm(y ~ x1 + x2, data = Morgage, family = binomial)

# Print the summary of the model
summary(Logit_mod)
```

## Compare these two models

Compare and interpret the estimation of the two models you developed above. 

```{r, echo=FALSE, message=FALSE, results='asis'}
library(stargazer)
stargazer(LPM_mod, Logit_mod, 
          type = "html")
```

## Predict a loan application

Predict the loan approval probability for an applicant with a 20% down payment and a 30% income-to-loan ratio using the two models respectively.

```{r, echo=FALSE}
# Predict using LPM
predict(LPM_mod, 
        newdata = data.frame(x1 = 20, x2 = 30), 
        type='response')
# Predict using logistic regression model
predict(Logit_mod, 
        newdata = data.frame(x1 = 20, x2 = 30), 
        type='response')
```

What if the down payment was 5%, income-to-loan ratio was 20%?

```{r, eval=FALSE, echo=FALSE}
# Predict using LPM
predict(LPM_mod, 
        newdata = data.frame(x1 = 5, x2 = 20), 
        type='response')
# Predict using logistic regression model
predict(Logit_mod, 
        newdata = data.frame(x1 = 5, x2 = 20), 
        type='response')
```




# Application 2: Spam detector

Electronic mail (email) is one of the most widely used Internet services.

While emails can greatly benefit productivity, they are susceptible to unsolicited, unwanted, and possibly virus-infested spam messages from illegitimate email addresses.

Peter Derby works as a cybersecurity analyst at a private equity firm.
His colleagues at the firm have been overwhelmed by large-scale spam emails.
Peter would like to implement an effective spam detection system on the company’s email server.
He has been tasked to create spam filters that detect spam emails and stop them from getting into email inboxes.

The objective is to detect spam based on the number of recipients, the number of hyperlinks, and the number of characters for each email.

## Download and import the [`Spam.csv`](examples/Spam.csv) dataset

```{r, message=FALSE}
library(readr)
Spam <- read_csv("examples/Spam.csv")
```

```{r, echo=FALSE}
library(kableExtra)
Spam %>%
  kbl() %>%
  kable_paper(full_width = F, position = "center") %>%
  scroll_box(width = "500px", height = "200px")
```

## Splitting the data into training and testing sets

```{r, eval=FALSE, echo=TRUE}
sample_index <- 
train <- 
test <- 
```

```{r, echo=FALSE}
sample_index <- sample(nrow(Spam),nrow(Spam)*0.80)
train <- Spam[sample_index,]
test <- Spam[-sample_index,]
```

## Fit the logistic regression model

Develop the logistic regression model for the Spam dataset. 

```{r, echo=FALSE, results='hide'}
# Assuming your dataset is named 'Spam'
Logit_mod <- glm(Spam ~ Recipients + Hyperlinks + Characters,
                 data = train, 
                 family = binomial)

# Print the summary of the model
summary(Logit_mod)
```

If you cannot write your own codes, try to ask EducateUsGPT assistant to help you out. 

```{r, echo=FALSE, results='asis'}
# Me: What question you should ask?
htmltools::includeHTML("questions/Q-323-Logit-spam.html")
```

## Making predictions

```{r}
train_pred <- predict(Logit_mod, newdata=train, type='response')
test_pred <- predict(Logit_mod, newdata=test, type='response')
```

## Evaluating the model

```{r, echo=FALSE}
# How should I evaluate a logistic regression model?
htmltools::includeHTML("questions/Q-125-purchase-evaluate.html")
```

```{r}
# Evaluating the accuracy
train_accuracy <- mean((train_pred > 0.5) == train$Spam)
print(paste('Train_Accuracy:', train_accuracy))
test_accuracy <- mean((test_pred > 0.5) == test$Spam)
print(paste('Test_Accuracy:', test_accuracy))

# Calculating AUC for training and testing sets

library(pROC)
train_auc <- roc(train$Spam, train_pred)
test_auc <- roc(test$Spam, test_pred)
train_auc; test_auc
```

[go to top](#header)
